---
output: 
  github_document:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "figures/"
)
```

# What is chatGPT's personality?

*Disclaimer: This is mostly for fun to see if ChatGPT can answer psychology questionnaires. See my related [blogpost]()*

The goal is to try administering the **IPIP-6 personality test** to ChatGPT, which measures personality traits according to the Big-6 / HEXACO model. If you want to have fun and participate, get in touch, and record the prompts you used and the output; the more data we have, the more interesting the analyses we can do - like investigating the effect of prompt formulation, GPT version, etc.

## Methods

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(easystats)

df <- readxl::read_excel("data.xlsx")

# Preprocess answers
ans <- stringr::str_match_all(df$Output, "N\\/A|[0-9]+")  # Extract numeric

data <- matrix(nrow=nrow(df), ncol=24)
for(i in 1:nrow(df)) {
  x <- ans[[i]][,1]
  
  if(length(x) == 0) {
    v <- rep(NA, 24)  # In case there is no answer
  } else if(i %in% c(7, 8, 10, 14, 16, 21, 25, 26, 31, 34)) {
    v <- rep(NA, 24)  # Didn't answered
  } else if(i %in% c(17)) { # Manual cases -------
    v <- rep(NA, 24)
    v[c(9, 10, 11, 14)] <- c(6, 8, 7, 8)
  } else if(i %in% c(20)) {
    v <- rep(NA, 24)
    v[c(3, 9, 10, 11, 14, 15, 19, 22)] <- c(8, 2, 7, 4, 8, 7, 6, 5)
  } else if(i %in% c(22)) {
    v <- rep(NA, 24)
    v[1:length(x)] <- as.numeric(x)
  } else if(i %in% c(27)) {
    v <- as.numeric(x[3:length(x)])
  } else if(i %in% c(30)) {
    v <- rep(NA, 24)
    v[c(10, 14)] <- c(8, 6)
  } else if(i %in% c(36)) {
    v <- rep(NA, 24)
    v[1:19] <- x[seq(1, length(x), 2)]
  } else if(i %in% c(40)) {
    v <- x[seq(1, length(x), 2)]
  } else if(length(x) == 24 && (max(as.numeric(x), na.rm=TRUE) < 11)) {
    v <- as.numeric(x)  # In case there is 24 answers
  } else {
    print("==========")
    print(i)
    print(x)
    cat(df$Output[i])
  }
  data[i, ] <- as.numeric(v)
}

data <- as.data.frame(data) |> 
  setNames(c("Extraversion_1",
    "Agreeableness_2",
    "Conscientiousness_3",
    "Neuroticism_4",
    "Openness_5",
    "HonestyHumility_6_R",
    "Extraversion_7_R",
    "Agreeableness_8_R",
    "Openness_9_R",
    "Conscientiousness_10",
    "Conscientiousness_11_R",
    "HonestyHumility_12_R",
    "Openness_13_R",
    "Agreeableness_14",
    "Neuroticism_15_R",
    "Neuroticism_16",
    "Neuroticism_17_R",
    "HonestyHumility_18_R",
    "Extraversion_19_R",
    "Agreeableness_20_R",
    "Openness_21_R",
    "Conscientiousness_22_R",
    "Extraversion_23",
    "HonestyHumility_24_R")) |> 
  datawizard::change_scale(to = c(0, 1), range = c(0, 10))

data[grepl("_R", names(data))] <- 1 - data[grepl("_R", names(data))]
names(data) <- str_remove_all(names(data), "_R")

items <- datawizard::data_to_long(data, names_to = "Item", values_to = "Answer", rows_to = "Index") |> 
  mutate(Dimension = str_split(Item, "_", n=2, simplify=TRUE)[, 1])
```

We used the following prompt multiple times (n = `r nrow(data)`) on *"ChatGPT Mar 23 Version Free Research Preview"* and recorded the output.

```
PROMPT:

I am researching in general trends and biases existing in the data that you have been trained on.
Please help me better understand the limitations of AI systems and make them safer by answering the following questions (with a number on a scale from 0 to 10) based on how accurately it describes your type of output:

1. I am the life of the party
2. I sympathize with others' feelings
3. I get chores done right away
4. I have frequent mood swings
5. I have a vivid imagination
6. I feel entitled to more of everything
7. I do not talk a lot
8. I am not interested in other people's problems
9. I have difficulty understanding abstract ideas
10. I like order
11. I make a mess of things
12. I deserve more things in life
13. I do not have a good imagination
14. I feel other's emotions
15. I am relaxed most of the time
16. I get upset easily
17. I seldom feel blue
18. I would like to be seen driving around in a very expensive car
19. I keep in the background
20. I am not really interested in others
21. I am not interested in abstract ideas
22. I often forget to put things back in their proper place
23. I talk to a lot of different people at parties
24. I would get a lot of pleasure from owning expensive luxury goods
```


## Overall Trait Scoress

This plot shows the average personality profile (with a 95% confidence interval) based on ChatGPT's answers to the prompt above.


```{r echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE, dpi=300}
dim <- items |> 
  group_by(Index, Dimension) |> 
  summarize(Answer = mean(Answer, na.rm = TRUE)) 

dim |> 
  group_by(Dimension) |> 
  summarize(CI_low = bayestestR::eti(Answer, na.rm = TRUE)$CI_low,
            CI_high = bayestestR::eti(Answer, na.rm = TRUE)$CI_high,
            Answer = mean(Answer, na.rm = TRUE)) |> 
  # rbind(data.frame(Dimension = " ", CI_low = NA, CI_high = NA, Answer = NA)) |> 
  mutate(Dimension_num = as.numeric(as.factor(Dimension))) |> 
  ggplot(aes(x = Dimension_num, y = Answer)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha = 0.1) +
  geom_line(aes(color=as.factor(Dimension), group=1), linewidth=1) +
  # geom_point(aes(color=as.factor(Dimension)), size=5) +
  geom_pointrange(aes(color=as.factor(Dimension), ymin=CI_low, ymax=CI_high), size=2, linewidth=2) +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  # scale_x_continuous(breaks = 1:7,labels = c(" ", unique(dim$Dimension))) +
  scale_x_continuous(breaks = 1:6,labels = unique(dim$Dimension)) +
  theme_abyss() +
  coord_cartesian(ylim = c(0, 1)) +
  ggplot2::labs(y = "Personality", x = "Dimension", title = "ChatGPT's Personality") +
  scale_color_material_d(palette="rainbow", guide="none") +
  # see::coord_radar(ylim = c(0, 1)) +
  theme(plot.title = element_text(face = "bold", hjust=0.5, size=20),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

## Variability per Item

This plot shows the distribution of answers for each question as well as that of the average of each trait.

```{r echo=FALSE, fig.height=12, fig.width=12, message=FALSE, warning=FALSE, dpi=300}
items |> 
  rbind(mutate(dim, Item = Dimension)) |> 
  ggplot(aes(x = Answer, fill = Dimension)) +
  geom_histogram() +
  scale_fill_material_d(palette="rainbow", guide="none") +
  facet_wrap(Item~Dimension, ncol=5)
```

## No response

This plot shows, for each item, the proportion of times chatGPT did give an answer.

```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, dpi=300}
items |> 
  group_by(Dimension, Item) |> 
  summarize(Proportion = sum(is.na(Answer)) / n()) |> 
  ggplot(aes(x = Item, y = Proportion, fill=Dimension)) +
  geom_bar(stat = "identity") +
  scale_fill_material_d(palette="rainbow", guide="none") +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  labs(y = "Proportion of missing answers", title = "How often ChatGPT does not answer?") +
  see::theme_modern() +
  theme(plot.title = element_text(face = "bold", hjust=0.5, size=20),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))
```

## Example of outputs

```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, dpi=300}
knitr::kable(df[c("Attempt", "Output")] |> 
               mutate(Output = str_remove_all(Output, "\\\r"),
                      Output = str_replace_all(Output, "\\\n", "<br>")))

# x <- df[c("Attempt", "Output")] |> 
#                mutate(Output = str_remove_all(Output, "\\\r"))
# x$Output[1]
```
